---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

```{r setup, include=FALSE, error=TRUE, message=FALSE}
# Set knitr options for knitting code into the report:
# - Don't print out code (echo)
# - Save results so that code blocks aren't re-run unless code changes (cache),
# _or_ a relevant earlier code block changed (autodep), but don't re-run if the
# only thing that changed was the comments (cache.comments)
# - Don't clutter R output with messages or warnings (message, warning)
  # This _will_ leave error messages showing up in the knitted report

knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE,
                      autodep = TRUE,
                      cache.comments = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      error = TRUE,
                      comment = NA)
```


```{r}
library(RcppOctave)

.CallOctave("version")

.CallOctave("sqrt", 10)

.CallOctave("eye", 3)

.CallOctave("eye", 3, 2)

```


```
% single output variable: eigen values only
S = svd(A);

% 3 output variables: complete SVD decomposition
[U, S, V] = svd(A);
```
Examples in ex_function.m belong to RcppOctave package.
```{r}
# source example function definitions from RcppOctave installation
sourceExamples("ex_functions.m")

# several functions are now defined
o_ls()
```

```{r}
# single output value
.CallOctave("fun1")
```

```{r}
# 3 output values
.CallOctave("fun2")
```

```{r}
# no output value
.CallOctave("fun_noargout", 1)

## % Printed from Octave: x= 1

.CallOctave("fun_noargout", "abc")

# variable number of arguments
.CallOctave("fun_varargin")


.CallOctave("fun_varargin", 1, 2, 3)
```

```{r}
.CallOctave("fun_varargout")
```

```{r}
.CallOctave("fun_varargout", argout = 1)
```


```{r}
# single output variable: result is S
.CallOctave("svd", matrix(1:4, 2))
```

```{r}
# 3 output variables: results is [U,S,V]
.CallOctave("svd", matrix(1:4, 2), argout = 3)
```

```{r}
# specify output names (and therefore number of output variables)
.CallOctave("svd", matrix(1:4, 2), argout = c("U", "S", "V"))
```

```{r}
.O

## <Octave Interface>
## - Use `$x` to call Octave function or get variable x.
## - Use `$x <- val` to assign a value val to the Octave variable x.

.O$version()

## [1] "3.6.4"

.O$eye(3)


.O$svd(matrix(1:4, 2))

# argout can still be specified
.O$svd(matrix(1:4, 2), argout = 3)
```

## Manipulating variables

```{r}
# define a variable
.O$myvar <- 1:5

# retrieve value
.O$myvar
```

```{r}
# assign and retrieve new value
.O$myvar <- 10
.O$myvar
```

```{r}
# remove
.O$myvar <- NULL

# this should now throw an error since 'myvar' does not exist anymore
try(.O$myvar)
```

## Calling functions

```{r}
# density of x=5 for Poisson(2)
.O$poisspdf(5, 2)

# E.g. compare with R own function
dpois(5, 2)
```

```{r}
# retrieve Octave function
f <- .O$poisspdf
f

## <OctaveFunction::`poisspdf`>

# call (in Octave)
f(5, 2)
```


```{r}
## ASSIGN
o_assign(a = 1)
o_assign(a = 10, b = 20)
o_assign(list(a = 5, b = 6, aaa = 7, aab = list(1, 2, 3)))

```


```{r}
## GET get all variables
str(o_get())
```

```{r}
# selected variables
o_get("a")

## [1] 5

o_get("a", "b")


# rename on the fly
o_get(c = "a", d = "b")
```

```{r}
# get a function
f <- o_get("svd")
f
```

## Evaluate single statements

```{r}
# assign variable 'a'
o_eval("a=1")
```

```{r}
o_eval("a") # or .O£a
```

```{r}
o_eval("a=svd(rand(3))")
```

```{r}
.O$a
```


```{r}
# eval a list of statements
l <- o_eval("a=rand(1, 2)", "b=randn(1, 2)", "rand(1, 3)")
l
```

```{r}
# variables 'a' and 'b' were assigned the new values
identical(list(.O$a, .O$b), l[1:2])
```

```{r}
# multiple statements are not supported by o_eval
try(o_eval("a=1; b=2"))
```

```{r}
.O$a
```

```{r}
# argument CATCH allows for recovering from errors in statement
o_eval("a=usage('ERROR: stop here')", CATCH = "c=3")
```

```{r}
# argument CATCH allows for recovering from errors in statement
o_eval("a=print_usage('ERROR: stop here')", CATCH = "c=3")
```

```{r}
.O$a
```

```{r}
.O$c
```

## Source .m files

```{r}
# clear all session
o_clear(all = TRUE)
o_ls()

## character(0)

# source example file from RcppOctave installation
mfile <- system.file("scripts/ex_source.m", package = "RcppOctave")
cat(readLines(mfile), sep = "\n")
```

```{r}
o_source(mfile)
```


```{r}
# Now objects 'a', 'b', and 'c' as well as the function 'abc' should be defined:
o_ls(long = TRUE)
```

```{r}
#
o_eval("abc(2, 4, 6)")
```

```{r}
o_eval("abc(a, b, c)")
```

```{r}
o_source(text = "clear a b c; a=100; a*sin(123)")

# last statement is stored in automatic variable 'ans'
o_get("a", "ans")
```

## List objects

```{r}
o_ls()
```
```{r}
o_ls(long = TRUE)
```

```{r}
# clear all (variables + functions)
o_clear(all = TRUE)
o_ls()
```

## Help

```{r}
o_help(std)
```


## Errors

```{r}
# error
res <- try(.CallOctave("error", "this is an error in Octave"))
```


```{r}
geterrmessage()
```

```{r}
# warning
res <- .CallOctave("warning", "this is a warning in Octave")
res
```

## Comparing R and Octave functions

```{r}
o_svd <- function(x) {
  # ask for the complete decomposition
  res <- .O$svd(x, argout = c("u", "d", "v"))
  # reformat/reorder result
  res$d <- diag(res$d)
  res[c(2, 1, 3)]
}

# define random data
X <- matrix(runif(25), 5)

# run SVD in R
svd.R <- svd(X)

# run SVD in Octave
svd.O <- o_svd(X)
str(svd.O)
```

```{r}
# check results
all.equal(svd.R, svd.O)
```

```{r}
# but not exactly identical
all.equal(svd.R, svd.O, tol = 10^-16)
```

## Random computations
In order to ensure reproducibility of results and facilitate the comparability of implementations between R and Octave , RcppOctave ships a custom Octave module that redeﬁne Octave standard random number generator functions rand, randn, rande and randg, so that they call R corresponding functions runif, rnorm, rexp and rgamma. This module is loaded when the RcppOctave package is itself loaded. As a result, random computation – that use these functions – can be seeded in both Octave and R, using R standard function set.seed. This facilitates, in particular, the validation of ports of stochastic algorithms (e.g. simulations, MCMC-based estimations):

```{r}
Rf <- function() {
  x <- matrix(runif(100), 10)
  y <- matrix(rnorm(100), 10)
  
  (x * y) %*% (x / y)
}

Of <- {
# define Octave function
o_source(text="
  function [res] = test()
    x = rand(10);
    y = randn(10);
    res = (x .* y) * (x ./ y);
  end
")
  
# return the function
.O$test
}

# run both computations with a common seed
set.seed(1234); res.R <- Rf()
set.seed(1234); res.O <- Of()

# compare results
identical(res.R, res.O)
```

```{r}
# not seeding the second computation would give different results
set.seed(1234);
identical(Rf(), Of())
```

