---
title: "MRST Example 03 book - pure Octave run"
output: html_notebook
---

## Introduction
This notebook shows how to set up core functions in MRST to run a very basic reservoir simulation. It is based on examples in the book "Geometric Modelling, Numerical Simulation, and Optimization" by Geir Hasle, Knut-Andreas Lie, Ewald Quak.

The Matlab scripts below are stand-alone which means you don't need to load anything else to run a basic reservoir simulation. At this stage, there is no need to install MRST. The scripts described below are enough.

```{r}
library(RcppOctave)
.CallOctave("version")
```

## Load the Matlab scripts in Octave
This work has been performed on a Linux Virtual Machine based on Mint 18.2 and Octave 4.0.


```{r}
library(RcppOctave)

o.TPFA <- OctaveFunction("
function [P,V] = TPFA(Grid, K, q)
    % Compute transmissibilities by harmonic averaging.
    Nx=Grid.Nx; Ny=Grid.Ny; Nz=Grid.Nz; N=Nx*Ny*Nz;
    
    hx=Grid.hx; hy=Grid.hy; hz=Grid.hz;
    L = K.^(-1);
    tx = 2*hy*hz/hx; TX = zeros(Nx+1,Ny,Nz);
    ty = 2*hx*hz/hy; TY = zeros(Nx,Ny+1,Nz);
    tz = 2*hx*hy/hz; TZ = zeros(Nx,Ny,Nz+1);
    TX(2:Nx,:,:) = tx./(L(1,1:Nx-1,:,:) + L(1,2:Nx ,:,:));
    TY(:,2:Ny,:) = ty./(L (2,:,1: Ny-1,:) + L(2,:,2:Ny,:));
    TZ (:,:,2: Nz) = tz./(L (3,:,:,1: Nz-1)+L(3,:,:,2:Nz));

    % Assemble TPFA discretization matrix.
    x1 = reshape(TX(1:Nx,:,:),N,1); x2 = reshape(TX(2:Nx+1,:,:),N,1);
    y1 = reshape(TY(:,1:Ny,:),N,1); y2 = reshape(TY(:,2:Ny+1,:),N,1);
    z1 = reshape(TZ(:,:,1:Nz),N,1); z2 = reshape(TZ(:,:,2:Nz+1),N,1);
    DiagVecs = [-z2,-y2,-x2,x1+x2+y1+y2+z1+z2,-x1,-y1,-z1];
    DiagIndx = [-Nx*Ny,-Nx,-1,0,1,Nx,Nx*Ny];
    
    A = spdiags(DiagVecs,DiagIndx,N,N);
    A(1,1) = A(1,1)+sum(Grid.K(:,1,1,1));
    
    % Solve linear system and extract interface ﬂuxes.
    u = A\\q;
    
    P = reshape(u,Nx,Ny,Nz);
    V.x = zeros(Nx+1,Ny,Nz);
    V.y = zeros(Nx,Ny+1,Nz);
    V.z = zeros(Nx,Ny,Nz+1);
    V.x(2:Nx ,:,:) = (P(1:Nx-1,:,:)-P(2:Nx,:,:)).*TX(2:Nx,:,:);
    V.y (:,2:Ny,:) = (P(:,1:Ny-1,:)-P(:,2:Ny,:)).*TY(:,2:Ny,:);
    V.z (:,:,2: Nz) = (P (:,:,1: Nz-1)-P(:,:,2:Nz)).*TZ(:,:,2:Nz);
")


o.GenA <- OctaveFunction("
function A = GenA(Grid,V,q)
    Nx=Grid.Nx; Ny=Grid.Ny; Nz=Grid.Nz; N=Nx*Ny*Nz;
    N=Nx*Ny*Nz; % number of unknowns
    
    fp=min(q,0); % production
    
    XN=min(V.x,0); x1=reshape(XN(1:Nx,:,:),N,1); % separate ﬂux into
    YN=min(V.y,0); y1=reshape(YN(:,1:Ny,:),N,1); % − ﬂow in positive coordinate
    
    ZN=min(V.z,0); z1=reshape(ZN(:,:,1:Nz),N,1); % direction (XP,YP,ZP)
    XP=max(V.x,0); x2=reshape(XP(2:Nx+1,:,:),N,1); % − ﬂow in negative coordinate
    
    YP=max(V.y,0); y2=reshape(YP(:,2:Ny+1,:),N,1); % direction (XN,YN,ZN)
    ZP=max(V.z,0); z2=reshape(ZP(:,:,2:Nz+1),N,1); %
    DiagVecs=[z2,y2,x2,fp+x1-x2+y1-y2+z1-z2,-x1,-y1,-z1]; % diagonal vectors
    DiagIndx=[-Nx*Ny,-Nx,-1,0,1,Nx,Nx*Ny]; % diagonal index
    
    A=spdiags(DiagVecs,DiagIndx,N,N); % matrix with upwind FV stencil
")




o.RelPerm <- OctaveFunction("
function [Mw,Mo,dMw,dMo] = RelPerm(s,Fluid)
    S = (s-Fluid.swc)/(1-Fluid.swc-Fluid.sor); % Rescale saturations
    
    Mw = S.^2/Fluid.vw; % Water mobility
    Mo =(1-S).^2/Fluid.vo; % Oil mobility
    if (nargout==4)
    dMw = 2*S/Fluid.vw/(1-Fluid.swc-Fluid.sor);
    dMo = -2*(1-S)/Fluid.vo/(1-Fluid.swc-Fluid.sor);
end                          
")



o.Pres <- OctaveFunction("
function [P,V] = Pres(Grid,S,Fluid,q)
    % Compute K∗lambda(S)
    
    [Mw,Mo] = RelPerm(S,Fluid);
    Mt = Mw+Mo;
    KM = reshape([Mt,Mt,Mt]',3,Grid.Nx,Grid.Ny,Grid.Nz).*Grid.K;
    % Compute pressure and extract ﬂuxes
    [P,V] = TPFA(Grid,KM,q);
")


o.Upstream <- OctaveFunction("
function S = Upstream(Grid,S,Fluid,V,q,T)

    Nx=Grid.Nx; Ny=Grid.Ny; Nz=Grid.Nz; % number of grid points
    N=Nx*Ny*Nz; % number of unknowns
    pv = Grid.V(:).*Grid.por(:); % pore volume=cell volume∗porosity
    fi =max(q,0); % inﬂow from wells
    XP=max(V.x,0); XN=min(V.x,0); % inﬂux and outﬂux, x−faces
    YP=max(V.y,0); YN=min(V.y,0); % inﬂux and outﬂux, y−faces
    ZP=max(V.z,0); ZN=min(V.z,0); % inﬂux and outﬂux, z−faces
    Vi = XP(1:Nx,:,:)+YP(:,1:Ny,:)+ZP(:,:,1:Nz)-... % total ﬂux into
    XN(2:Nx+1,:,:)-YN(:,2:Ny+1,:)-ZN(:,:,2:Nz+1); % each gridblock
    
    pm = min(pv./(Vi(:)+fi)); % estimate of inﬂux
    cfl = ((1-Fluid.swc-Fluid.sor)/3)*pm; % CFL restriction
    
    Nts = ceil(T/cfl); % number of local time steps
    dtx = (T/Nts)./pv; % local time steps
    
    A = GenA(Grid,V,q); % system matrix
    A=spdiags(dtx,0,N,N)*A; % A ∗ dt/|Omega i|
    fi =max(q,0).*dtx; % injection
    for t=1:Nts
    [mw,mo]=RelPerm(S,Fluid); % compute mobilities
    fw = mw./(mw+mo); % compute fractional ﬂow
    S = S+(A*fw+fi); % update saturation
    
    end                           
")
```


## Run the example
All the scripts are practically executed in Octave. The graphics window is also opened by sending the plot commands `contourf` and `drawnow`.

```{r}
o.runExample <- OctaveFunction("
function [Grid, N, S, Fluid, Q] = runExample()

    Grid.Nx=64; Dx=1; Grid.hx = Dx/Grid.Nx; % Dimension in x−direction
    Grid.Ny=64; Dy=1; Grid.hy = Dy/Grid.Ny; % Dimension in y−direction
    Grid.Nz=1; Dz=1; Grid.hz = Dz/Grid.Nz; % Dimension in z−direction
    N = Grid.Nx*Grid.Ny;                    % Total number of grid blocks
    Grid.V = Grid.hx*Grid.hy*Grid.hz;        % Cell volumes
     
    Grid.K = ones(3,Grid.Nx,Grid.Ny,Grid.Nz); % Unit permeability
    Grid.por = ones(Grid.Nx,Grid.Ny,Grid.Nz); % Unit porosity
    Q = zeros(N,1); Q([1 N])=[1 -1]; % Production/injection
    Fluid.vw = 1.0; Fluid.vo=1.0; % Viscosities
    Fluid.swc = 0.0; Fluid.sor=0.0; % Irreducible saturations

    S = zeros(N,1);             % Initial saturation

    nt = 25; dt = 0.7/nt; % Time steps
    for t=1:nt
        [P,V]=Pres(Grid,S,Fluid,Q); % pressure solver
        S=Upstream(Grid,S,Fluid,V,Q,dt); % saturation solver
       % plot ﬁlled contours at the midpoints of the grid cells
       contourf(linspace(Grid.hx/2,Dx-Grid.hx/2,Grid.Nx),...
       linspace(Grid.hy/2,Dy-Grid.hy/2,Grid.Ny),...
       reshape(S,Grid.Nx,Grid.Ny),11,'k');
       axis square; caxis([0 1]); % equal axes and color
       drawnow; % force update of plot
    end
")

runExample <- o.runExample()

```
