---
title: "R Notebook"
output: html_notebook
---

```{r}
library(R6)

# an empty class
Grid <- R6Class("Grid")

# instantiate class
g <- Grid$new()
g
```

```{r}
Grid = R6Class("Grid", 
               public = list( 
                   initialize = function(name) { 
                       print(paste("A new Grid with name '", name, "'")) 
                   }
                   )
                 )

g1 <- Grid
g1
```

```{r}
# iniialize the class with a name
g2 <- Grid$new("grid 2")
```

## Putting the variable in a field
```{r}
Grid = R6Class("Grid",
     public = list(
       Name = NA,
       initialize = function(name) 
           {self$Name = name }
       )
)
# new instance
g = Grid$new("Grid #1")
paste("This grid is called", g$Name)
# change it
g$Name = "Grid #2"
paste("This grid is now called", g$Name)
```

## Active bindings
If you want to have getter-setter method you need to use the so-called **active bindings**. This is a get-set just like in JavaScript which returns the field if no value is provided but sets the field if one is given. If the method has no argument it’s presumed a read-only property.

```{r}
Grid = R6Class("Grid",
                private = list(
                  name = NA
                  ),
                active = list(
                  Name = function(name){
                    if(missing(name)) return(private$name)
                    private$name = name
                  }
                  ),
                public = list(                  
                   initialize = function(name) {
                       private$name = name }
                   )
)
# new instance
g = Grid$new("Grid #1")
paste("This grid is called", g$Name)
# change it
g$Name = "Grid #2"
paste("This grid is now called", g$Name)

```


## Adding methods
```{r}
Grid = R6Class("Grid",
                private = list(
                  name = NA,
                  speakInternal = function() {
                      print("A private message.")}
                  ),
                active = list(
                  Name = function(name){
                    if(missing(name)) return(private$name)
                    private$name = name
                  }
                  ),
                public = list(                  
                   initialize = function(name){private$name = name },
                   Speak = function(){ 
                       print("A standard grid.")}
                   )
                 )
g = Grid$new("grid #3")
g$Speak()
g$.__enclos_env__$private$speakInternal()
```

## Inheritance

```{r}
ComplexGrid <- R6Class("ComplexGrid",
                       inherit = Grid
                       )

cg <- ComplexGrid$new("a complex grid")
cg$Speak()
```

## Calling the parent method
```{r}
ComplexGrid = R6Class("ComplexGrid",
               inherit = Grid,
               public = list(
                 initialize = function(){
                   super$initialize("another complex grid")
                 }
                 )
               )

cg <- ComplexGrid$new()

cg$Speak()
```

## 
```{r}
ComplexGrid = R6Class("ComplexGrid",
               inherit = Grid,
               public = list(
                 initialize = function(){
                   super$initialize("A complex grid")
                 },
                 Speak = function(){
                                    print("This is a complex grid even if the base-class says,")
                                    super$Speak()
                                    }
                 )
               )
cg <- ComplexGrid$new()
cg$Speak()
```

## Adding methods on the fly

```{r}
# cannot run this lines twice
ComplexGrid$set("public", "Grid30", function(){"30x30 grid"})

# Error in ComplexGrid$set("public", "Grid30", function() { : Can't add Grid30 because it already present in ComplexGrid generator.

cg2 <- ComplexGrid$new()
cg2$Grid30()
```


```{r}
Reservoir = R6Class("Reservoir", public = list(
    Orient = Grid$new("Grid-East")
  ))

reservoir1 = Reservoir$new()
print(paste("The orientation in reservoir 1 is ", reservoir1$Orient$Name))
reservoir1 = Reservoir$new()
print(paste("The orientation in reservoir 2 is also", reservoir1$Orient$Name))
```

```{r}
Reservoir = R6Class("Reservoir", public = list(
    Orient = NA,
    initialize = function(name){self$Orient =  Grid$new(name)  }
  ))

reservoir1 = Reservoir$new("West")
print(paste("The orientation in reservoir 1 is", reservoir1$Orient$Name))

reservoir2 = Reservoir$new("South")
print(paste("The orientation in reservoir 2 is", reservoir2$Orient$Name))
```

Source: http://www.orbifold.net/default/2015/04/24/r6-classes/

# Another example

```{r}
Grid <- R6Class("Grid",
  public = list(
    x = NULL,
    initialize = function(x = 1) self$x <- x,
    getx = function() self$x,
    inc  = function(n = 1) self$x <- self$x + n
  )
)

g <- Grid$new()
g$getx()
g$inc()
g$getx()

g$inc(10)
g$getx()
```

## Private and public members

```{r}
GridPrivate <- R6Class("GridPrivate",
  private = list(x = NULL),
  public = list(
    initialize = function(x = 1) private$x <- x,
    getx = function() private$x,
    inc = function(n = 1) private$x <- private$x + n
  )
)

gp <- GridPrivate$new()
gp$getx()
gp$inc()
gp$getx()
```

## Actve Bindings

```{r}
Grid.Nx =8; Grid.hx=1/Grid.Nx;
Grid.Ny =8; Grid.hy=1/Grid.Ny;
Grid.Nz =1; Grid.hz=1/Grid.Nz;

Grid.K = m.ones(3, Grid.Nx, Grid.Ny);
```


```{r}
library(rMRST)

Grid <- R6Class("Grid",
    public = list(
        Nx = 0, Ny = 0, Nz = 0,
        hx = 0, hy = 0, hz = 0,
        K  = NA,
        N  = NA,
    initialize = function(Nx, Ny, Nz) { 
        self$Nx = Nx
        self$Ny = Ny
        self$Nz = Nz
        self$hx = 1 / self$Nx
        self$hy = 1 / self$Ny
        self$hz = 1 / self$Nz
        self$K  = array(1, c(3, self$Nx, self$Ny))
        self$N  = self$Nx * self$Ny * self$Nz
        cat(sprintf("Grid of %dx%dx%d", self$Nx, self$Ny, self$Nz))
        image(self$K[1,,])
        }
    )
)

grid <- Grid$new(8, 8, 1)
# Grid of 8x8x1
```

Source: https://rpubs.com/wch/24456

```{r}
Numbers <- R6Class("Numbers",
  public = list(
    x = 100
  ),
  active = list(
    x2 = function(value) {
      if (missing(value)) return(self$x * 2)
      else self$x <- value/2
    },
    rand = function() rnorm(1)
  )
)

n <- Numbers$new()
n$x
```



